---
title: "The Leslie Matrix - revisited"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

A population of fish we're interested in has four life stages: eggs (E), fry (F), juvenile (J), breeding adult (A). You are told that annually:

-   Each breeding adult will produce, on average, 600 eggs
-   5% of eggs will survive to become fry
-   20% of fry will survive to become juveniles
-   16% of juveniles will survive to become adults
-   Adult survival rate year to year is 90%

Recall the Leslie matrix from EDS 212. Matrix multiplication (using the dot product `%*%` of the matrix with the populations at each life stage) yields the population structure at the following year.

### Create the Leslie matrix:

```{r}
# Make the Leslie Matrix:
fish_leslie <- matrix(c(0, 0, 0, 600, 0.05, 0, 0, 0, 0, 0.20, 0, 0, 0, 0, 0.16, 0.9), nrow = 4, ncol = 4, byrow = TRUE)

# Check it out: 
fish_leslie
```

The initial population structure is given by:

```{r}
# Initial population structure (Roe, Fry, Juvenile, Adult): 
fish_year0 <- c(5000, 8000, 600, 400)
# take dot product of matrix and this pop structure of fish_year0
```

Add line-by-line comments wheres prompted in the code below, explaining what each is doing. Make sure that you are talking with your neighbors about what is really being created at each step.

```{r}
# Model projections

# ------ WHAT HAPPENS HERE? ------ #
# ------ comment line-by-line ---- #

# store a sequence for time (sequence from o to 8 by 1)
time_yr <- seq(from = 0, to = 8, by = 1)

# create an empty matrix with 4 columns and 9 rows (0 to 8) [9, 4]
proj_year <- matrix(ncol = length(fish_year0), nrow = length(time_yr))

# We already have the numbers for fish_year0 (putting those values in row 1)
# "I want row 1 of proj_year to be populated by fish_year0 values
proj_year[1, ] <- fish_year0

# don't use seq_along bc I don't want to populate things in row 1 (already populated)
# take dot product of the fish_leslie matrix and the row containing the population structure data from the previous cycle 
# saying that we want this information to populate the next row
for (i in 2:length(time_yr)) {
  proj_year[i,] <- fish_leslie %*% proj_year[i-1,]
}


# The rest is wrangling & visualization (run to check it out):
# reassigning column names 
colnames(proj_year) <- c("eggs", "fry", "juvenile", "adult")

# binding vector of times with different life stage populations
# pivot_longer across all columns except time_year (could have also just written cols = c("eggs", "fry", "juvenile", "adult"))
# take all column names and put them into a single columns
# take all values and put them correlating to life stages in another column
proj_df <- data.frame(time_yr, proj_year) %>% 
  pivot_longer(cols = -time_yr, names_to = "lifestage", values_to = "stage_population")
# creating a plot - scale_y_log10() is used to scale it
ggplot(data = proj_df, aes(x = time_yr, y = stage_population)) +
  geom_line(aes(color = lifestage)) +
  scale_y_log10()

```

Test different combinations of survival / reproduction and starting population structures to see what happens!

```{r}
```
